---
title: "Function of Summarizing Groups"
author: "Ivy Liu"
date: "`r Sys.Date()`"
output: html_document
Roxygen: list(markdown = TRUE)
---

```{r setup, include=FALSE}
library(dplyr)
library(testthat)
```

## Exercise 1: Make a Function (25 points)
## Exercise 2: Document your Function (20 points)
In this exercise, we will make a function that take a data frame or a tibble, and output the summary of data after grouping.

```{r}
#' @title Summary of Groups
#' @details This function output a summary table after grouping of your data.
#' @param data the dataset you want to do analysis on
#' @param group_cols a vector of column names where the grouping occurs
#' @param summary_cols a vector of variable names to perform statistical summary upon
#' @param summary_functions a list summarizing functions
#' @return returns a summary table with the output of the summarizing functions after grouping

group_summary <- function(data, group_cols, summary_cols, summary_functions) {
  for(i in summary_cols){
    if(any(is.na(data[which(names(data) == i)]))){
      stop("Sorry, this function only works for data with no NA entry.")
    }
  }
  data %>%
    group_by(across({{group_cols}})) %>%
    summarize(across(all_of({{summary_cols}}),
                     .fns = summary_functions,
                     .names = "{.col}_{.fn}"),
              .groups = "keep")
}
```

## Exercise 3: Include examples (15 points)
We begin with a simple example where the function only compute the mean after grouping.
```{r}
mydata1 = tibble(group = c("A", "A", "B", "B"),
                number = c(1,2,3,4))
group_summary(mydata1, group, summary_cols = "number", list(Mean = mean))
```
The output is what we expected because it correctly computes the mean of number in group A and group B.

Now we move on to a slightly more complicated example, where we perform grouping by two categorical variables, and perform two summarizing functions on two numerical variables.
```{r}
mydata2 = tibble(group1 = c("A", "A", "A", "A", "B", "B", "B", "B"),
                 group2 = c("C", "C", "D", "D", "C", "C", "D", "D"),
                 number1= c(5,5,6,6,7,7,8,8),
                 number2 = c(1, 1, 2, 2, 3, 3, 4, 4))
group_summary(data = mydata2, group_col = c(group1, group2), 
              summary_cols = c("number1", "number2"), 
              list(Mean = mean, Median = median))
```
The result is also expected. The function first splits the data into four groups (AC, AD, BC, and BD), and then compute the mean and median of both number1 and number2 in each group.

There is one drawback of the customized function: this function cannot perform the count function "n()" as a summary because it defaultly passes the colnames in "summary_cols" into the summary function, i.e. "n(number1)", which is an invalid syntax.
```{r, error = TRUE}
group_summary(data = mydata2, group_col = c(group1), 
              summary_cols = c("number1"), 
              list(Count = n))
```



## Exercise 4: Test the Function (25 points)
```{r}
# test of a normal dataset
test_data = tibble(group1 = rep(c(1,2), each = 4),
                   value = rep(c(1,2), each = 4))
test_that("Testing the numeric output", {
  expect_equal(group_summary(data = test_data, group_col = c(group1), 
              summary_cols = c("value"), 
              list(Mean = mean))$value_Mean, c(1,2) )
  expect_equal(group_summary(data = test_data, group_col = c(group1), 
              summary_cols = c("value"), 
              list(Mean = mean))$group1, c(1,2))
})
```

```{r}
# test of a dataset with NA's
test_data_na = tibble(group1 = rep(c(1,2), each = 4),
                   value = rep(c(1, NA, 2, NA), each = 2))
test_that("Testing the numeric output", {
  expect_error(group_summary(data = test_data_na, group_col = c(group1), 
              summary_cols = c("value"), list(Mean = mean)), 
              "Sorry, this function only works for data with no NA entry.")
})
```

```{r}
# test of a multiple grouping on a data frame type of input
test_data_2_grouping = data.frame(group1 = rep(c(1,2), each = 4),
                                  group2 = rep(c(3,4), times = 4),
                                  value = c(1,2,3,4,5,6,7,8))
test_that("Testing multiple grouping", {
  expect_equal(colnames(group_summary(data = test_data_2_grouping, 
                             group_col = c(group1,group2), 
                             summary_cols = c("value"), 
                             list(Mean = mean, Median = median))), 
               c("group1", "group2", "value_Mean", "value_Median"))
  expect_equal(group_summary(data = test_data_2_grouping, 
                             group_col = c(group1,group2), 
                             summary_cols = c("value"), 
                             list(Mean = mean, Median = median))$value_Mean, 
               c(2, 3, 6, 7))
  expect_equal(group_summary(data = test_data_2_grouping, 
                             group_col = c(group1,group2), 
                             summary_cols = c("value"), 
                             list(Mean = mean, Median = median))$value_Median, 
               c(2, 3, 6, 7))
})
```



